---
title: 'Lab: Class One Survey Analysis'
output: github_document
---

### **INSTRUCTIONS**

#### **Complete all tasks 1-7 using reproducible coding practices (annotation of all code). Use R syntax to solve all problems (i.e. do not manually change values/variables in datasets or check them manually to make sure things are correct except as a double or triple check). Please cut and paste each problem before each R markdown chunk and annotate your code within the chunks.**

##### 1. Import class one survey data from our Github site (<https://github.com/kijohnson/Advanced-Data-Analysis>) (The dataset is called 'Class 1 Survey Fall 2026.csv'), calling the R dataframe that you create *C1survey*.

```{r}
# Import the CSV file from GitHub and save it as C1survey
C1survey <- read_csv("https://raw.githubusercontent.com/kijohnson/Advanced-Data-Analysis/refs/heads/main/Class%201%20Survey%20Fall%202026.csv")
```

##### 2. Determine: **a.** the number of observations (i.e. the number of people who filled out the survey) and **b.** the number of variables in the dataframe.

```{r}

# Check the dimensions of C1 survey (n=31)
dim(C1survey)
```

##### 3. **a.** Rename the column variables to something shorter and that is descriptive of what the variable is about (for example *like_dogs* for the 'Do you like dogs?' question variable) and **b.** Display that they are renamed. **Hints**: You can use either the rename_with function or the names function. The list of names for time is:

```{r, echo=TRUE}
#Get column variable names

c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")
```

```{r}

# Rename columns 1-27 of C1survey with the new titles
names(C1survey)[1:27] <-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")

```

##### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset. **Suggestion**: Try to ask ChatGPT for a solution if you are unsure. There is a really short answer to this but we want to see what students come up with!

```{r}

# Use the sum function to count the number of factor, integer, numeric, and character variables and print below


sum(sapply(C1survey, is.factor))
# 0 factor variables

sum(sapply(C1survey, is.integer))
# 0 integer variables

sum(sapply(C1survey, is.numeric))
# 4 numeric variables

sum(sapply(C1survey, is.character)) 
# 23 character variables
```

##### 5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values (**hint**: use table function). If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. \*Note: please make sure that your bday and month variables are numeric before doing statistics

```{r}

# Make a table of birth dates and birth months
table(C1survey$bday, C1survey$bmonth)

# There were two birthdates with the months as a string and a few birth months as strings instead of numeric. I would recode these to all be numeric.

C1survey$bday[C1survey$bday == "30-Nov"] <- "30"
C1survey$bday[C1survey$bday == "29-Nov"] <- "29"
C1survey$bmonth[C1survey$bmonth == "May"] <- "5"
C1survey$bmonth[C1survey$bmonth == "March"] <- "3"
C1survey$bmonth[C1survey$bmonth == "November"] <- "11"
C1survey$bmonth[C1survey$bmonth == "February"] <- "2"

# Make an updated table to check
table(C1survey$bday, C1survey$bmonth)
```

##### 6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.

```{r}

# Get the class of bmonth (character)
class(C1survey$bmonth)

# Change birth month to be numeric and store under month_numeric
month_numeric <- as.numeric(C1survey$bmonth)

# Check the class of month_numeric to make sure
class(month_numeric)

# Make an array called bseason
C1survey$bseason <- array(C1survey, dim = c(31))

# Omit missing birth month values
month_numeric <- na.omit(month_numeric)

# Make an array to go through each value (person) of bseason, and for each person, use logic to determine the season
for (i in 1:length(C1survey$bseason)) {
  if ((month_numeric[i] == 12) || (month_numeric[i] <= 2)) {
    C1survey$bseason[i] = "winter"
  } else if ((month_numeric[i] >= 3) && (month_numeric[i] <= 5)) {
    C1survey$bseason[i] = "spring"
  } else if ((month_numeric[i] >= 6) && (month_numeric[i] <= 8)) {
    C1survey$bseason[i] = "summer"
  } else {
    C1survey$bseason[i] = "fall"
  }
}

```

##### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. \#'s).

```{r}

# Question: How many people had the same favorite season as the season they were born in?

# Make a new variable for favorite season so the code isn't overwritten
C1survey$season_lowercase <- C1survey$fav_season

# Make the favorite seasons lowercase to match the birth seasons
C1survey$season_lowercase <- tolower(C1survey$season_lowercase)


# Check if favorite season is equal to birth season, and if so, print the word "True"

for (i in 1:length(C1survey$season_lowercase)) {
  if (C1survey$season_lowercase[i] == C1survey$bseason[i]) {
      print ("True")
  }
}

# Output: 10 "True"s, meaning 10 people had the same favorite season and birth season- can verify this with the data
```

```         
```
